A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$

### 2025-04-08 paper_v02/article.tex
conclusion

The $66.50\%$ increase in performance from the basic to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations. Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale. These results are presented in Fig.~\ref{fig:elapsed_tws_04}.

@@

The \pkg{Zebende} package implementations where designed to increase performance in scenarios with large number of points in the series, but also with a large amount of time series used in the $DMC_x^2$ calculations. To prove that, the same tests were also made using the same data, with the combinations of twelve series, 2 by 2. The rasults of that second test are presented in Table~\ref{tab:time_12}.

@@

With that configuration, the optimized \proglang{Zig} version still the fastest, followed by the basic \proglang{Zig} implementation. The \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages. The lists ended with the \pkg{DFA} in the 6th position.

@@

@@

The \pkg{Zebende} package is the first tool that calculates the $DMC_{x}^{2}$ coefficient for any number of time series. The package also calculates the $\rho_{DCCA}$, a fundamental step for the $DMC_{x}^{2}$ calculation. The package was tested with data from the \cite{RIBEIRO2025107419} paper, and the results were validated. The package was also compared with other packages that calculate the $\rho_{DCCA}$, and the results showed that the \pkg{Zebende} package is the fastest. The package is available at the \textbf{pypi}.

>>>>

The $66.50\%$ increase in performance in the basic compared to the optimized \proglang{Zig} implementation is a good result, showcasing the effectiveness of the optimizations. Another test was conducted, wth the same data and settings as the previous one, but getting the elapsed time inside every time scale. These results are presented in Fig.~\ref{fig:elapsed_tws_04}.

@@

The \pkg{Zebende} package implementations where designed to increase performance in scenarios with large number of points in the series, but also with a large amount of time series used in the $DMC_x^2$ calculations. To prove that, the same tests were also made using the same data, with the combinations of twelve series, 2 by 2. The results of that second test are presented in Table~\ref{tab:time_12}.

@@

@@

With twelve time series, the optimized \proglang{Zig} version still the fastest, followed by the basic \proglang{Zig} implementation. The \pkg{Zebende} pure \proglang{Python} implementation ended in the third position, performing better than the \pkg{DCCA-Python} (4th) and the \pkg{DCCA-R-Java} (5th) packages. The lists ended with the \pkg{DFA} in the 6th position.

Comparing the basic with the optimized \proglang{Zig} algorithms using twelve time series the performance increase is $76.64\%$. Figure \ref{fig:elapsed_tws_12} shows the comparison between the basic and optimized \proglang{Zig} implementations with twelve time series. The results are equivalent to the ones obtained with four time series, reinforcing the capabilities of the optimization technics.

The codes used for testing are available in \url{replace}, as jupyter noteooks, in  and can easily be tested in Google colab environment. The basic version is available in \pkg{Zebende} version $0.1.91$ and the optimized implementation in version $0.2.0$ of the package.

@@

The \pkg{Zebende} package is the first tool that calculates the $DMC_{x}^{2}$ coefficient for any number of time series. The package also calculates the $DFA$, $DCCA$ and the $\rho_{DCCA}$, a fundamental step for the $DMC_{x}^{2}$ calculation. The package was tested with data from the \cite{RIBEIRO2025107419} paper, and the results were validated. The package was also compared with other packages that calculate the $\rho_{DCCA}$, and the results showed that the \pkg{Zebende} package is the fastest, making it suitable for large datasets.

The tests with twelve time scales also shows that the \pkg{Zebene} implementation, designed to obtain better performance working with greater number of time series, deliver great results even with the pure \proglang{Python} non-optimized version. This package enable the exploration of the $DMC_x^2$ coefficient in scenarios where it was never tested before. The package is available at the \textbf{pypi}.
